---
title: "R Notebook"
output: html_notebook
---
#gts need to have at least till 2034
```{r}
library(knitr)
library(tidyverse)
library(stringr)
library(here)
library(fs)
library(readxl)
library(dplyr)
```

Load vars
```{r}

source(here("scripts", "set_vars.R"))

country_ref <- read_rds(here("C:/R projects/ist_2024_08/output_data/country_ref.rds"))


```

#the latest gts is in gct folder
```{r}
data_1 <- read_csv(path(path_te_shared, path_gts_current, file_gts_current))

data_2 <- data_1 %>%
  filter(location=="Canada")

```

#adding spending and nights
```{r}

data_3_re <- data_2 %>% 
  mutate(variable = case_when(indicator == "Arrivals" ~ "v",
                              indicator == "Overnight tourists arrivals" ~ "v",
                              indicator == "Nights" ~ "n",
                              str_detect(indicator_code, "FNIGHTSTOT") ~ "n",
                              indicator == "Spending" ~ "s",
                              T ~ NA_character_)) %>%
  # set total
  mutate(origin_iso3166_1a3 = case_when(indicator == "Overnight tourists arrivals" ~ "WLD",
                                        indicator == "Nights in all accommodation types (international)" ~ "WLD",
                                        indicator == "Spending" & origin == "World" ~ "WLD",
                                 T ~ origin_iso3166_1a3)) 


```

#if the orig_code is NA, we still keep the orig_ttl
#data_6 has all regions, countries, including NES("Not elsewhere specified") and WLD("Total international")
#the orig_ttl is not yet standardized here bc regions don't have orig_code so it will leads to multiple matching
```{r}
data_6 <- data_3_re %>%
	     select(loc_ttl=location,
	            loc_code=location_code, 
	            date=year,
	            orig_code=origin_iso3166_1a3,
	            orig_ttl=origin,
	            variable,
	            indicator, 
	            indicator_code, 
	            value) %>%
	     filter(variable %in% c("v", "s", "n")) %>%
       mutate(orig_code = ifelse(orig_ttl == "Not elsewhere specified", "NES", orig_code)) %>%
	     mutate(orig_ttl = ifelse(orig_code == "WLD" & !is.na(orig_code), "Total international", orig_ttl)) %>%
	     mutate(date = as.Date(paste0(date, "-01-01")))

colSums(is.na(data_6))
```


#data_4_ctry has all countries
```{r}
data_4_ctry <- data_3_re %>%
  select(loc_ttl=location,
         loc_code=location_code, 
         date=year,
         orig_code=origin_iso3166_1a3,
         orig_ttl=origin,
         variable,
         value) %>%
  filter(variable %in% c("v", "s", "n") & !is.na(orig_code)) %>%
  mutate(orig_ttl = ifelse(orig_code=="WLD", "Total international", orig_ttl)) %>%
  mutate(date = as.Date(paste0(date, "-01-01"))) %>%
  filter(orig_ttl != "Total international") %>%
  left_join(country_ref %>% select(iso3166_1a3, country_name), by = c("orig_code" = "iso3166_1a3")) %>%
  select(-c(orig_ttl)) %>%
  rename(orig_ttl = country_name)

colSums(is.na(data_4_ctry))
```
#data_4_ctry_NES has all countries and NES(37 years so this dataframe has 74 more obs than data_4_ctry, arrivals and nights)
#orig_ttl is the standadized name
```{r}
data_4_ctry_NES <- data_3_re %>%
  select(loc_ttl=location,
         loc_code=location_code, 
         date=year,
         orig_code=origin_iso3166_1a3,
         orig_ttl=origin,
         variable,
         value) %>%
  filter(variable %in% c("v", "s", "n") & (!is.na(orig_code) | orig_ttl == "Not elsewhere specified")) %>%
  mutate(orig_code = ifelse(orig_ttl == "Not elsewhere specified", "NES", orig_code)) %>%
  mutate(orig_ttl = ifelse(orig_code=="WLD", "Total international", orig_ttl)) %>%
  mutate(date = as.Date(paste0(date, "-01-01"))) %>%
  filter(orig_code != "WLD") %>%
  left_join(country_ref %>% select(iso3166_1a3, country_name), by = c("orig_code" = "iso3166_1a3")) %>%
  select(-c(orig_ttl)) %>%
  rename(orig_ttl = country_name) %>%
  mutate(orig_ttl = ifelse(orig_code == "NES", "Not elsewhere specified", orig_ttl))
  

colSums(is.na(data_4_ctry_NES))
```
#Burundi(BDI), Chad(TCD), Gabon(GAB), Papua New Guinea(PNG), Uzbekistan(UZB),NES don't exist in spending, but exist in arrivals and nights. That's 6*37 year = 222 obs less in spending.
```{r}
mod_frame_data_4_ctry_NES <- apply(subset(data_4_ctry_NES, select = -c(value)), 2 , table)
print(mod_frame_data_4_ctry_NES)
```



```{r}
gts_current_ctry_NES <- data_4_ctry_NES
write_rds(gts_current_ctry_NES, here("output_data", "gts_current_ctry_NES.rds"))

# gts_current_reg_ctry <- data_6
# write_rds(gts_current_reg_ctry, here("output_data", "gts_current_reg_ctry.rds"))
# write_csv(gts_current_reg_ctry, here(path(path_te_shared, path_ist_est_2408), "gts_current_reg_ctry.csv"))

```

