---
title: "R Notebook"
output: html_notebook
---

```{r}
library(dplyr)
library(knitr)
library(tidyverse)
library(stringr)
library(here)
library(fs)
library(readxl)

source(here("scripts", "set_vars.R"))
```

```{r}
gts_current_ctry_NES <- read_rds(here("output_data", "gts_current_ctry_NES.rds")) #current gts for Canada
ctry_tour <- read_rds(here("output_data", "ctry_tour.rds"))
reg_tour <- read_rds(here("output_data", "reg_tour.rds"))

reg_def <- read_rds(here("output_data", "reg_def.rds"))

non_us_data3 <- read_rds(here("output_data", "non_us_data3.rds"))
```

```{r}
gts_v <- gts_current_ctry_NES %>%
  filter(variable == "v" ) %>%
  mutate(value = 1000*value)

```


```{r}
can_ctry_visit <- ctry_tour %>%
  filter(geo == "Canada") %>%
  left_join(select(gts_v, orig_code, date, gts_v = value), by = c("orig_code", "date")) %>%
  mutate(ratio = value/gts_v)

#2022 and before all align with gts visits
test <- can_ctry_visit %>%
  filter(date <= "2022-01-01")

#to lower gts future to match stat_can 2023 level
ratio_2023 <- can_ctry_visit %>%
  filter(date == "2023-01-01") %>%
  select(orig_code, ratio)
```

#see if sum of provinces is the same as total Canada
#some are not really equal to total Canada
```{r}
test1 <- ctry_tour %>%
  filter(geo != "Canada") %>%
  group_by(orig_code, date) %>%
  summarise(sum = sum(value, na.rm = TRUE), .groups = "drop") %>%
  ungroup() %>%
  left_join(select(can_ctry_visit, orig_code, date, can_v = value), by = c("orig_code", "date")) %>%
  mutate(gap = sum-can_v) %>%
  mutate(geo = "Canada")
```

#make history sum of states be Canada
```{r}
df1 <- ctry_tour %>%
  left_join(select(test1, geo, date, orig_code, sum), by = c("geo", "date", "orig_code")) %>%
  mutate(value = ifelse(is.na(sum), value, sum)) %>%
  select(-c(sum)) %>%
  tidyr::complete(geo, date, orig_code, fill = list(value = 0))

colSums(is.na(df1))
```
#calculate province share of Canada for each country
```{r}
pro_share <- df1 %>%
  filter(date == "2019-01-01" | date == "2023-01-01") %>%
  group_by(orig_code, geo) %>%
  summarise(value = sum(value, na.rm = TRUE), .groups = "drop") %>%
  ungroup() %>%
  group_by(orig_code) %>%
  mutate(avg_share = value / value[geo == "Canada"]) %>%
  ungroup()
```

```{r}
pro_share_complete <- df1 %>%
  group_by(orig_code, date) %>%
  mutate(share_c = value / value[geo == "Canada"]) %>%
  ungroup()

share_forecast <- pro_share_complete %>%
  left_join(select(pro_share, orig_code, geo, avg_share), by = c("orig_code", "geo")) %>%
  filter(date == "2023-01-01") %>%
  mutate(share_2024 = 0.3*avg_share + 0.7*share_c,
         share_2025 = 0.3*avg_share + 0.7*share_2024,
         share_2026 = 0.3*avg_share + 0.7*share_2025,
         share_2027 = 0.3*avg_share + 0.7*share_2026,
         share_2028 = 0.3*avg_share + 0.7*share_2027,
         share_2029 = 0.3*avg_share + 0.7*share_2028,
         share_2030 = 0.3*avg_share + 0.7*share_2029,
         share_2031 = 0.3*avg_share + 0.7*share_2030,
         share_2032 = 0.3*avg_share + 0.7*share_2031,
         share_2033 = 0.3*avg_share + 0.7*share_2032,
         share_2034 = 0.3*avg_share + 0.7*share_2033)

share_forecast2 <- share_forecast %>%
  select(geo, orig_code, share_2024:share_2034) %>%
  pivot_longer(cols = share_2024:share_2034, names_to = "date", values_to = "share_future") %>%
  mutate(date = substr(date, nchar(date)-3, nchar(date))) %>%
  mutate(date = as.Date(paste0(date, "-01-01")))
```


#complete the date
```{r}
df2 <- df1 %>%
  select(geo, date, orig_code, country_of_residence, low_lev_reg, traveller_type) %>%
  complete(nesting(geo, orig_code, country_of_residence, low_lev_reg, traveller_type), date = seq.Date(from=as.Date("2008-01-01"), to=as.Date("2034-01-01"), by="year"))
```

#construct country final dataframe
```{r}
df3 <- df2 %>%
  left_join(df1, by = c("geo", "date", "orig_code", "country_of_residence", "low_lev_reg", "traveller_type")) %>%  #history
  left_join(select(gts_v, date, orig_code, gts_v = value), by = c("date", "orig_code")) %>%
  left_join(ratio_2023, by = c("orig_code")) %>%
  left_join(select(share_forecast2, orig_code, geo, date, share_future), by = c("orig_code", "date", "geo")) %>%
  mutate(value = case_when(date <= "2023-01-01" ~ value,
                           date >= "2024-01-01" ~ gts_v * ratio * share_future)) %>%
  select(-c(gts_v, ratio, share_future))

colSums(is.na(df3))
```

#Apply gts growth rate on statcan region
```{r}
ctry_def2 <- data.frame(orig_code= c("GRL", "MEX", "SPM", "BLZ", "CRI", "SLV", "GTM", "HND", "NIC", "PAN", "AIA", "ATG", "ABW", "BHS", "BRB", "BMU", "BES", "CYM", "CUB", "CUW", "DMA", "DOM", "GRD", "GLP", "HTI", "JAM", "MTQ", "MSR", "ANT", "PRI", "BLM", "KNA", "LCA", "MAF", "VCT", "SXM", "TTO", "TCA", "VGB", "VIR", "ARG", "BOL", "BRA", "CHL", "COL", "ECU", "FLK", "GUF", "GUY", "PRY", "PER", "SGS", "SUR", "URY", "VEN", "AUT", "BEL", "FRA", "DEU", "LIE", "LUX", "MCO", "NLD", "CHE", "ARM", "AZE", "BLR", "BGR", "CZE", "EST", "GEO", "HUN", "KAZ", "KGZ", "LVA", "LTU", "MDA", "POL", "ROU", "RUS", "SVK", "TJK", "TKM", "UKR", "UZB", "ALA", "DNK", "FRO", "FIN", "GGY", "ISL", "IRL", "IMN", "JEY", "NOR", "", "SJM", "SWE", "GBR", "ALB", "AND", "BIH", "HRV", "GIB", "GRC", "VAT", "ITA", "", "MLT", "MNE", "MKD", "PRT", "SMR", "SRB", "SVN", "ESP", "CYP", "ISR", "TUR", "BEN", "BFA", "CPV", "CIV", "GMB", "GHA", "GIN", "GNB", "LBR", 
                                     "MLI", "MRT", "NER", "NGA", "SHN", "SEN", "SLE", "TGO", "IOT", "BDI", "COM", "DJI", "ERI", "ETH", "KEN", "MDG", "MWI", "MUS", "MYT", "MOZ", "REU", "RWA", "SYC", "SOM", "TZA", "UGA", "ZMB", "ZWE", "DZA", "MAR", "SSD", "SDN", "TUN", "ESH", "AGO", "CMR", "CAF", "TCD", "COD", "COG", "GNQ", "GAB", "STP", "BWA", "SWZ", "LSO", "NAM", "ZAF", "BHR", "EGY", "IRQ", "JOR", "KWT", "LBN", "LBY", "OMN", "QAT", "SAU", "SYR", "ARE", "", "YEM", "CHN", "HKG", "JPN", "PRK", "KOR", "MAC", "MNG", "TWN", "BRN", "KHM", "IDN", "LAO", "MYS", "MMR", "PHL", "SGP", "THA", "TLS", "VNM", "AFG", "BGD", "BTN", "IND", "IRN", "MDV", "NPL", "PAK", "LKA", "AUS", "NZL", "FJI", "NCL", "NFK", "PNG", "SLB", "VUT", "CXR", 
                                     "CCK", "GUM", "KIR", "MHL", "FSM", "NRU", "MNP", "PLW", "UMI", "ASM", "COK", "PYF", "NIU", "PCN", "WSM", "TKL", "TON", "TUV", "WLF", "ATA", "BVT", "ATF", "HMD"),
                       low_lev_reg=c("North America, countries other than the United States of America", "North America, countries other than the United States of America", "North America, countries other than the United States of America", "Central America", "Central America", "Central America", "Central America", "Central America", "Central America", "Central America", "Caribbean", "Caribbean", "Caribbean", "Caribbean", "Caribbean", "Caribbean", "Caribbean", "Caribbean", "Caribbean", 
                                     "Caribbean", "Caribbean", "Caribbean", "Caribbean", "Caribbean", "Caribbean", "Caribbean", "Caribbean", "Caribbean", "Caribbean", "Caribbean", "Caribbean", "Caribbean", "Caribbean", "Caribbean", "Caribbean", "Caribbean", "Caribbean", "Caribbean", "Caribbean", "Caribbean", "South America", "South America", "South America", "South America", "South America", "South America", "South America", "South America", "South America", "South America", "South America", 
                                     "South America", "South America", "South America", "South America", "Western Europe", "Western Europe", "Western Europe", "Western Europe", "Western Europe", "Western Europe", "Western Europe", "Western Europe", "Western Europe", "Central/Eastern Europe", "Central/Eastern Europe", "Central/Eastern Europe", "Central/Eastern Europe", "Central/Eastern Europe", "Central/Eastern Europe", "Central/Eastern Europe", "Central/Eastern Europe", "Central/Eastern Europe", "Central/Eastern Europe", "Central/Eastern Europe", "Central/Eastern Europe", "Central/Eastern Europe", "Central/Eastern Europe", "Central/Eastern Europe", "Central/Eastern Europe", "Central/Eastern Europe", 
                                     "Central/Eastern Europe", "Central/Eastern Europe", "Central/Eastern Europe", "Central/Eastern Europe", "Northern Europe", "Northern Europe", "Northern Europe", "Northern Europe", "Northern Europe", "Northern Europe", "Northern Europe", "Northern Europe", "Northern Europe", "Northern Europe", "Northern Europe", "Northern Europe", "Northern Europe", "Northern Europe", "Southern Europe", "Southern Europe", "Southern Europe", "Southern Europe", 
                                     "Southern Europe", "Southern Europe", "Southern Europe", "Southern Europe", "Southern Europe", "Southern Europe", "Southern Europe", "Southern Europe", "Southern Europe", "Southern Europe", "Southern Europe", "Southern Europe", "Southern Europe", "East Mediterranean Europe", "East Mediterranean Europe", "East Mediterranean Europe", "Western Africa", "Western Africa", "Western Africa", "Western Africa", "Western Africa", "Western Africa", "Western Africa", "Western Africa", "Western Africa", "Western Africa", "Western Africa", "Western Africa", "Western Africa", "Western Africa", "Western Africa", "Western Africa", "Western Africa", "Eastern Africa", "Eastern Africa", 
                                     "Eastern Africa", "Eastern Africa", "Eastern Africa", "Eastern Africa", "Eastern Africa", "Eastern Africa", "Eastern Africa", "Eastern Africa", "Eastern Africa", "Eastern Africa", "Eastern Africa", "Eastern Africa", "Eastern Africa", "Eastern Africa", "Eastern Africa", "Eastern Africa", "Eastern Africa", "Eastern Africa", "Northern Africa", "Northern Africa", "Northern Africa", "Northern Africa", "Northern Africa", "Northern Africa", "Central Africa", 
                                     "Central Africa", "Central Africa", "Central Africa", "Central Africa", "Central Africa", "Central Africa", "Central Africa", "Central Africa", "Southern Africa", "Southern Africa", "Southern Africa", "Southern Africa", "Southern Africa", "Middle East", "Middle East", "Middle East", "Middle East", "Middle East", "Middle East", "Middle East", "Middle East", "Middle East", "Middle East", "Middle East", "Middle East", "Middle East", "Middle East", 
                                     "Northeast Asia", "Northeast Asia", "Northeast Asia", "Northeast Asia", "Northeast Asia", "Northeast Asia", "Northeast Asia", "Northeast Asia", "Southeast Asia", "Southeast Asia", "Southeast Asia", "Southeast Asia", "Southeast Asia", "Southeast Asia", "Southeast Asia", "Southeast Asia", "Southeast Asia", "Southeast Asia", "Southeast Asia", "Southern Asia", "Southern Asia", "Southern Asia", "Southern Asia", "Southern Asia", "Southern Asia", "Southern Asia", 
                                     "Southern Asia", "Southern Asia", "Australasia", "Australasia", "Melanesia", "Melanesia", "Melanesia", "Melanesia", "Melanesia", "Melanesia", "Micronesia", "Micronesia", "Micronesia", "Micronesia", "Micronesia", "Micronesia", "Micronesia", "Micronesia", "Micronesia", "Micronesia", "Polynesia", "Polynesia", "Polynesia", "Polynesia", "Polynesia", "Polynesia", "Polynesia", "Polynesia", "Polynesia", "Polynesia", "Antarctica and Adjacent Islands Region", 
                                     "Antarctica and Adjacent Islands Region", "Antarctica and Adjacent Islands Region", "Antarctica and Adjacent Islands Region"))
```

#Manually give Micronesia and Polynisia estimated growth
#Manually add Antarctica
```{r}
gts_v_rest_reg <- gts_v %>%
  left_join(ctry_def2, by = c("orig_code")) %>%
  filter(!orig_code %in% unique(df3$orig_code)) %>%
  group_by(date, low_lev_reg) %>%
  summarise(sum_rest_reg = sum(value, na.rm = TRUE), .groups = "drop") %>%
  ungroup() %>%
  group_by(low_lev_reg) %>%
  mutate(growth = sum_rest_reg/sum_rest_reg[date == "2023-01-01"]) %>%
  ungroup()

oce <- gts_v_rest_reg %>%
  filter((low_lev_reg == "Polynesia" | low_lev_reg == "Micronesia"| low_lev_reg == "Melanesia") & date >= "2024-01-01") %>%
  group_by(date) %>%
  mutate(growth_oce = growth[low_lev_reg == "Melanesia"]) %>%
  ungroup() %>%
  select(-c(sum_rest_reg, growth))

ata <- gts_v_rest_reg %>%
  filter(low_lev_reg == "Melanesia") %>%
  mutate(low_lev_reg = "Antarctica and Adjacent Islands Region") %>%
  select(-c(sum_rest_reg))

nma <- gts_v_rest_reg %>%
  filter(low_lev_reg == "Western Europe") %>%
  mutate(low_lev_reg = "North America, countries other than the United States of America") %>%
  select(-c(sum_rest_reg))

complete_rest_growth <- gts_v_rest_reg %>%
  select(-c(sum_rest_reg)) %>%
  rbind(ata, nma) %>%
  left_join(oce, by = c("date", "low_lev_reg")) %>%
  mutate(growth = ifelse(is.na(growth_oce), growth, growth_oce)) %>%
  select(-c(growth_oce))
```

#complete the date
```{r}
df4 <- reg_tour %>%
  tidyr::complete(geo, date, country_of_residence, fill = list(value = 0)) %>%
  select(geo, date, country_of_residence, high_lev_reg, traveller_type) %>%
  complete(nesting(geo, country_of_residence, high_lev_reg, traveller_type), date = seq.Date(from=as.Date("2008-01-01"), to=as.Date("2034-01-01"), by="year"))
```


```{r}
#sum up countries within region
df3_reg <- df3 %>%
  filter(orig_code != "USA") %>%
  group_by(geo, low_lev_reg, date) %>%
  summarise(value_reg = sum(value, na.rm = TRUE), .groups = "drop") %>%
  ungroup()
```

#calculate low level rest of region
```{r}
df5 <- df4 %>%
  filter(!is.na(high_lev_reg)) %>%
  left_join(select(reg_tour, geo, date, country_of_residence, value), by = c("geo", "date", "country_of_residence")) %>%
  left_join(df3_reg, by = c("geo", "country_of_residence" = "low_lev_reg", "date")) %>%
  left_join(complete_rest_growth, by = c("date", "country_of_residence" = "low_lev_reg")) %>%
  mutate(value_reg = ifelse(is.na(value_reg),0,value_reg)) %>%
  mutate(value_rest = value - value_reg) %>%
  mutate(low_lev_reg = country_of_residence, .after = country_of_residence) %>%
  mutate(country_of_residence = paste0("Rest of ",country_of_residence)) %>%
  select(-c(value, value_reg)) %>%
  group_by(geo, country_of_residence) %>%
  mutate(value_rest = case_when(date >= "2024-01-01" ~ value_rest[date== "2023-01-01"] * growth,
                           TRUE ~ value_rest)) %>%
  ungroup() %>%
  mutate(value_rest = ifelse(is.na(value_rest), 0, value_rest)) %>%
  select(-c(growth)) %>%
  mutate(orig_code = NA, .after = geo) %>%
  rename(value = value_rest)
```


#calculate low level region
```{r}
low_lev <- df3 %>%
  left_join(reg_def, by = c("low_lev_reg")) %>%
  select(geo, orig_code, country_of_residence, low_lev_reg, high_lev_reg, traveller_type, date, value) %>%
  rbind(df5) %>%
  filter(!is.na(low_lev_reg) & country_of_residence != "United States of America residents entering Canada") %>%
  group_by(geo, low_lev_reg, high_lev_reg, date) %>%
  summarise(value = sum(value, na.rm = TRUE), .groups = "drop") %>%
  ungroup()
```

#calculate high level region
```{r}
high_lev <- low_lev %>%
  group_by(geo, date, high_lev_reg) %>%
  summarise(value = sum(value, na.rm = TRUE), .groups = "drop") %>%
  ungroup()
```

#calculate high level rest of region
```{r}
df6 <- df5 %>%
  group_by(geo, date, high_lev_reg, traveller_type) %>%
  summarise(value = sum(value, na.rm = TRUE), .groups = "drop") %>%
  ungroup() %>%
  mutate(high_lev_reg = paste0("Rest of ",high_lev_reg))
```

#calculate "Residents of countries other than the United States of America entering Canada"
```{r}
non_us_lev <- high_lev %>%
  group_by(geo, date) %>%
  summarise(value = sum(value, na.rm = TRUE), .groups = "drop") %>%
  ungroup() %>%
  mutate(country_of_residence = "Residents of countries other than the United States of America entering Canada", .after = date)
```


#compare if high level is the sum of low level in statcan
#tiny different
```{r}
test2 <- reg_tour %>%
  filter(is.na(high_lev_reg)) %>%
  left_join(high_lev, by = c("geo", "date", "country_of_residence" = "high_lev_reg")) %>%
  mutate(gap = value.x-value.y)
```

#compare if non US is the sum of high level in statcan
#tiny different
```{r}
test3 <- non_us_data3 %>%
  filter(year <= 2023) %>%
  group_by(year, geo) %>%
  summarise(value = sum(value, na.rm = TRUE), .groups = "drop") %>%
  ungroup() %>%
  mutate(date = as.Date(paste0(year, "-01-01"))) %>%
  left_join(non_us_lev, by = c("geo", "date")) %>%
  mutate(gap = value.x-value.y)
```



```{r}
ctry_v <- df3 %>%
  mutate(high_lev_reg = NA, .after = low_lev_reg)

low_lev_reg_v <- low_lev %>%
  mutate(orig_code = NA,
         country_of_residence = low_lev_reg,
         traveller_type = "Tourists (overnight)") %>%
  select(geo, orig_code, country_of_residence, low_lev_reg, high_lev_reg, traveller_type, date, value)

low_lev_reg_v_rest <- df5 %>%
  mutate(orig_code = NA,
         low_lev_reg = country_of_residence,
         traveller_type = "Tourists (overnight)") %>%
  select(geo, orig_code, country_of_residence, low_lev_reg, high_lev_reg, traveller_type, date, value)

high_lev_reg_v <- high_lev %>%
  mutate(orig_code = NA,
         country_of_residence = high_lev_reg,
         low_lev_reg = NA,
         traveller_type = "Tourists (overnight)") %>%
  select(geo, orig_code, country_of_residence, low_lev_reg, high_lev_reg, traveller_type, date, value)

high_lev_reg_v_rest <- df6 %>%
  mutate(orig_code = NA,
         country_of_residence = high_lev_reg,
         low_lev_reg = NA,
         traveller_type = "Tourists (overnight)") %>%
  select(geo, orig_code, country_of_residence, low_lev_reg, high_lev_reg, traveller_type, date, value)

non_US_v <- non_us_lev %>%
  mutate(orig_code = NA,
         low_lev_reg = NA,
         high_lev_reg = NA,
         traveller_type = "Tourists (overnight)") %>%
  select(geo, orig_code, country_of_residence, low_lev_reg, high_lev_reg, traveller_type, date, value)

ttl_int <- ctry_v %>%
  filter(orig_code == "USA") %>%
  rbind(non_US_v) %>%
  group_by(geo, date) %>%
  summarise(value = sum(value, na.rm = TRUE), .groups = "drop") %>%
  ungroup() %>%
  mutate(orig_code = NA, 
         country_of_residence = "Non-resident visitors entering Canada", 
         low_lev_reg = NA, 
         high_lev_reg = NA,  
         traveller_type = NA) %>%
  select(geo, orig_code, country_of_residence, low_lev_reg, high_lev_reg, traveller_type, date, value)
```

```{r}
final_1 <- ctry_v %>%
  rbind(low_lev_reg_v) %>%
  rbind(low_lev_reg_v_rest) %>%
  rbind(high_lev_reg_v) %>%
  rbind(high_lev_reg_v_rest) %>%
  rbind(non_US_v) %>%
  rbind(ttl_int)

```

#calculate Canada
```{r}
final_can <- final_1 %>%
  filter(geo != "Canada") %>%
  group_by(orig_code, country_of_residence, low_lev_reg, high_lev_reg, traveller_type, date) %>%
  summarise(value = sum(value, na.rm = TRUE),
            .groups = "drop") %>%
  ungroup() %>%
  mutate(geo = "Canada", .before = orig_code)
```

```{r}
final_df_v <- final_1 %>%
  filter(geo != "Canada") %>%
  rbind(final_can)
```

```{r}
colSums(is.na(ctry_v))
colSums(is.na(low_lev_reg_v))
colSums(is.na(high_lev_reg_v))
colSums(is.na(low_lev_reg_v_rest))
colSums(is.na(high_lev_reg_v_rest))
colSums(is.na(non_US_v))

write_rds(ctry_v, here("output_data", "ctry_v.rds"))
write_rds(low_lev_reg_v, here("output_data", "low_lev_reg_v.rds"))
write_rds(high_lev_reg_v, here("output_data", "high_lev_reg_v.rds"))
write_rds(low_lev_reg_v_rest, here("output_data", "low_lev_reg_v_rest.rds"))
write_rds(high_lev_reg_v_rest, here("output_data", "high_lev_reg_v_rest.rds"))
write_rds(non_US_v, here("output_data", "non_US_v.rds"))

# write.csv(ctry_v, here(path(path_te_shared, path_output), "ctry_v.csv"))
# write.csv(low_lev_reg_v, here(path(path_te_shared, path_output), "low_lev_reg_v.csv"))
# write.csv(high_lev_reg_v, here(path(path_te_shared, path_output), "high_lev_reg_v.csv"))
# write.csv(low_lev_reg_v_rest, here(path(path_te_shared, path_output), "low_lev_reg_v_rest.csv"))
# write.csv(high_lev_reg_v_rest, here(path(path_te_shared, path_output), "high_lev_reg_v_rest.csv"))
# write.csv(non_US_v, here(path(path_te_shared, path_output), "non_US_v.csv"))

write_rds(final_df_v, here("output_data", "final_df_v.rds"))
write.csv(final_df_v, here(path(path_te_shared, path_output), "final_df_v.csv"))

write_rds(ctry_def2, here("output_data", "ctry_def2.rds"))

```

