---
title: "R Notebook"
output: html_notebook
---

```{r}
library(dplyr)
library(knitr)
library(tidyverse)
library(stringr)
library(here)
library(fs)
library(readxl)

source(here("scripts", "set_vars.R"))
options(scipen=999)
```

```{r}
final_df_v<- read_rds(here("output_data", "final_df_v.rds"))
final_df_s<- read_rds(here("output_data", "final_df_s.rds"))
final_df_s_0919<- read_rds(here("output_data", "final_df_s_0919.rds"))


#source data-visits
ctry_tour <- read_rds(here("output_data", "ctry_tour.rds"))
reg_tour <- read_rds(here("output_data", "reg_tour.rds"))
non_us_data3 <- read_rds(here("output_data", "non_us_data3.rds"))

#source data-spending
spend_data3 <- read_rds(here("output_data", "spend_data3.rds"))
spend_pro_data3 <- read_rds(here("output_data", "spend_pro_data3.rds"))

#processed source data-spending
df_ctrl_ovs_f <- read_rds(here("output_data", "df_ctrl_ovs_f.rds"))
df_ctrl2_complete_f <- read_rds(here("output_data", "df_ctrl2_complete_f.rds"))

#current gts for Canada
gts_current_ctry_NES <- read_rds(here("output_data", "gts_current_ctry_NES.rds")) 

#reference
ctry_def2<- read_rds(here("output_data", "ctry_def2.rds"))
ctry_def3<- read_rds(here("output_data", "ctry_def3.rds"))
reg_def <- read_rds(here("output_data", "reg_def.rds"))
```

#check diff
```{r}
check_us <- final_df_s %>%
  left_join(final_df_s_0919, by = c("geo", "area_of_residence", "date")) %>%
  mutate(gap_v = v.x-v.y,
         gap_s = s_forecast.x-s_forecast.y)
```


#check sum of provinces and total Canada
```{r}
check_v_1 <- final_df_v %>%
  filter(geo != "Canada") %>%
  group_by(country_of_residence, date) %>%
  summarise(sum_v = sum(value, na.rm = TRUE), .groups = "drop") %>%
  ungroup() %>%
  left_join(final_df_v %>%filter(geo == "Canada"), by = c("country_of_residence", "date")) %>%
  mutate(gap_v = value - sum_v)
```

```{r}
check_s_1 <- final_df_s %>%
  filter(geo != "Canada") %>%
  group_by(area_of_residence, date) %>%
  summarise(sum_v = sum(v, na.rm = TRUE),
            sum_s = sum(s_forecast, na.rm = TRUE),
            .groups = "drop") %>%
  ungroup() %>%
  left_join(final_df_s %>%filter(geo == "Canada"), by = c("area_of_residence", "date")) %>%
  mutate(gap_v = v - sum_v,
         gap_s = s_forecast - sum_s)
```

#check with source data-visits
```{r}
#compare country visits to Canada and provinces
#the gap is from sum provinces and total Canada difference from raw data
check_v_2_ctry <- ctry_tour %>%
  left_join(select(final_df_v, geo, orig_code, date, v_cal = value), by = c("geo", "orig_code", "date")) %>%
  mutate(gap_v = value - v_cal)

#compare region visits to Canada and provinces
#sum of Yukon northern + central+... Africa is 88, not 69. So this gap is passed on to total Canada as well
#sum of Yukon Australasia + ... in Oceania is 1084, not 1803. So this gap is passed on to total Canada as well
check_v_2_reg <- reg_tour %>%
  left_join(select(final_df_v, geo, country_of_residence, date, v_cal = value), by = c("geo", "country_of_residence", "date")) %>%
  mutate(gap_v = value - v_cal)

check_v_2_ovs <- non_us_data3 %>%
  filter(year <= 2023) %>%
  mutate(date = as.Date(paste0(year, "-01-01"))) %>%
  group_by(date, geo, country_of_residence) %>%
  summarise(value = sum(value, na.rm = TRUE),
            .groups = "drop") %>%
  ungroup() %>%
  left_join(select(final_df_v, geo, country_of_residence, date, v_cal = value), by = c("geo", "country_of_residence", "date")) %>%
  mutate(gap_v = value - v_cal)
```

#check with source data-spending-1st dataset
```{r}
check_s_2_1 <- spend_data3 %>%
  filter(year <= "2023-01-01") %>%
  group_by(year, geo, area_of_residence) %>%
  summarise(value = sum(value, na.rm = TRUE), .groups = "drop") %>%
  ungroup() %>%
  mutate(date = as.Date(paste0(year, "-01-01"))) %>%
  left_join(final_df_s, by = c("geo", "area_of_residence", "date")) %>%
  mutate(gap_s = value - s_forecast)
#the big gap is from "Non-resident visitors to Canada", which in statcan includes same day US tourists as well. It makes sense it doesn't align
#other gaps comes from: statcan region is missing for some period while countries have numbers. I overwrite the region
#US overnight and countries aligns
```

#check with source data-spending-2st dataset
```{r}
check_s_2_2 <- spend_pro_data3 %>%
  filter(year <= "2023-01-01" & area_of_residence == "Residents of countries other than the United States of America") %>%
  group_by(year, geo, area_of_residence) %>%
  summarise(value = sum(value, na.rm = TRUE), .groups = "drop") %>%
  ungroup() %>%
  mutate(date = as.Date(paste0(year, "-01-01"))) %>%
  left_join(final_df_s, by = c("geo", "area_of_residence", "date")) %>%
  mutate(gap_s = value - s_forecast)
#let's say 2018 Canada overseas should be 11479945 (sum up everything under overseas in 1st dataset), 4 more than overseas in second dataset. I distributed the gap to each province.
```

#check with processed source data-spending-1st dataset
```{r}
check_s_2_1_processed <- df_ctrl2_complete_f %>%
  filter(date <= "2023-01-01") %>%
  left_join(final_df_s, by = c("geo", "area_of_residence", "date")) %>%
  mutate(gap_s = value - s_forecast)
```

#check with processed source data-spending-2st dataset
```{r}
check_s_2_2_processed <- df_ctrl_ovs_f %>%
  filter(date <= "2023-01-01" & area_of_residence == "Residents of countries other than the United States of America") %>%
  left_join(final_df_s, by = c("geo", "area_of_residence", "date")) %>%
  mutate(gap_s = value - s_forecast)
#let's say 2018 Canada overseas should be 11479945 (sum up everything under overseas in 1st dataset), 4 more than overseas in second dataset. I distributed the gap to each province.
```

```{r}
#make the region better
final_df_v_process1 <- final_df_v %>%
  mutate(low_lev_reg = case_when(str_detect(country_of_residence, 'Rest of') & !is.na(low_lev_reg) ~ gsub("Rest of ", "", low_lev_reg),
                                 TRUE ~ low_lev_reg),
         high_lev_reg = case_when(str_detect(country_of_residence, 'Rest of') & is.na(low_lev_reg) ~ gsub("Rest of ", "", high_lev_reg),
                                 TRUE ~ high_lev_reg)) %>%
  left_join(select(reg_def, low_lev_reg, join_reg = high_lev_reg), by = c("low_lev_reg")) %>%
  mutate(high_lev_reg = ifelse((str_detect(country_of_residence, 'Rest of') & is.na(low_lev_reg) | country_of_residence %in% unique(join_reg)), high_lev_reg, join_reg)) %>%
  mutate(low_lev_reg = ifelse(country_of_residence == "United States of America residents entering Canada", NA, low_lev_reg)) %>%
  select(-c(traveller_type, join_reg))

 v_ref_for_client <- final_df_v_process1 %>%
   select(orig_code, country_of_residence, low_lev_reg, high_lev_reg) %>%
   distinct()
```

#check sum of country and rest of region, and compare with low lev region
```{r}
check_v_3_1 <- final_df_v_process1 %>%
  filter(!(country_of_residence %in% unique(reg_def$low_lev_reg)) & !is.na(low_lev_reg)) %>%
  group_by(geo, low_lev_reg, date) %>%
  summarise(v_sum = sum(value, na.rm = TRUE), .groups = "drop") %>%
  ungroup() %>%
  left_join(select(final_df_v_process1, geo, country_of_residence, date, value), by = c("geo", "low_lev_reg" = "country_of_residence", "date")) %>%
  mutate(gap_v = value - v_sum)
```

#check sum of low lev reg, and compare with high lev region
```{r}
check_v_3_2 <- check_v_3_1 %>%
  select(-c(value, gap_v)) %>%
  left_join(reg_def, by = c("low_lev_reg")) %>%
  group_by(geo, high_lev_reg, date) %>%
  summarise(v_sum = sum(v_sum, na.rm = TRUE), .groups = "drop") %>%
  ungroup() %>%
  left_join(select(final_df_v_process1, geo, country_of_residence, date, value), by = c("geo", "high_lev_reg" = "country_of_residence", "date")) %>%
  mutate(gap_v = value - v_sum)
```

#check sum of high lev reg, and compare with overseas
```{r}
check_v_3_3 <- check_v_3_2 %>%
  select(-c(value, high_lev_reg, gap_v)) %>%
  group_by(geo, date) %>%
  summarise(v_sum = sum(v_sum, na.rm = TRUE), .groups = "drop") %>%
  ungroup() %>%
  mutate(country_of_residence = "Residents of countries other than the United States of America entering Canada") %>%
  left_join(select(final_df_v_process1, geo, country_of_residence, date, value), by = c("geo", "country_of_residence", "date")) %>%
  mutate(gap_v = value - v_sum)
```

#check sum of overseas and USA, and compare with total international
```{r}
check_v_3_4 <- final_df_v_process1 %>%
  filter(country_of_residence %in% c("Residents of countries other than the United States of America entering Canada", "United States of America residents entering Canada")) %>%
  group_by(geo, date) %>%
  summarise(v_sum = sum(value, na.rm = TRUE), .groups = "drop") %>%
  ungroup() %>%
  mutate(country_of_residence = "Non-resident visitors entering Canada") %>%
  left_join(select(final_df_v_process1, geo, country_of_residence, date, value), by = c("geo", "country_of_residence", "date")) %>%
  mutate(gap_v = value - v_sum)
```


#check sum of things in spending

```{r}
reg_def2 <- data.frame(s_ttl= c("Not elsewhere specified-country in Americas, countries other than the United States of America", "Americas, countries other than the United States of America - other countries", "Mexico", "Caribbean", "Brazil", "Europe - other countries", "Belgium", "France", "Germany", "Netherlands", "Switzerland", "Scandinavia", "United Kingdom", "Italy", "Not elsewhere specified-country in Europe", "Africa - other countries", "Northern Africa", "Not elsewhere specified-country in Africa", "Asia and Oceania - other countries", "China", "Hong Kong", "Japan", "Southern Asia", "India", "Australia", "New Zealand","Not elsewhere specified-country in Asia and Oceania", "Antarctica and Adjacent Islands - other countries", "Not elsewhere specified-country"),
                             s_high_lev_reg=c("Americas, countries other than the United States of America", "Americas, countries other than the United States of America", "Americas, countries other than the United States of America", "Americas, countries other than the United States of America", "Americas, countries other than the United States of America", "Europe", "Europe", "Europe", "Europe", "Europe", "Europe", "Europe", "Europe", "Europe", "Europe", "Africa", "Africa", "Africa", "Asia and Oceania", "Asia and Oceania", "Asia and Oceania", "Asia and Oceania", "Asia and Oceania", "Asia and Oceania", "Asia and Oceania", "Asia and Oceania", "Asia and Oceania", "Antarctica and Adjacent Islands", "Not elsewhere specified-region"))
```

#sum of country and rest of region, and compare with region
```{r}
check_s_3_1 <- final_df_s %>%
  left_join(reg_def2, by = c("area_of_residence" = "s_ttl")) %>%
  filter(!is.na(s_high_lev_reg)) %>%
  group_by(geo, s_high_lev_reg, date) %>%
  summarise(v_sum = sum(v, na.rm = TRUE),
            s_sum = sum(s_forecast, na.rm = TRUE), .groups = "drop") %>%
  ungroup() %>%
  left_join(select(final_df_s, geo, area_of_residence, date, v, s_forecast), by = c("geo", "s_high_lev_reg" = "area_of_residence", "date")) %>%
  mutate(gap_v = v - v_sum,
         gap_s = s_forecast - s_sum)
```

#sum of region, and compare with overseas
```{r}
check_s_3_2 <- check_s_3_1 %>%
  group_by(geo, date) %>%
  summarise(v_sum = sum(v_sum, na.rm = TRUE),
            s_sum = sum(s_sum, na.rm = TRUE), .groups = "drop") %>%
  mutate(area_of_residence = "Residents of countries other than the United States of America") %>%
  ungroup() %>%
  left_join(select(final_df_s, geo, area_of_residence, date, v, s_forecast), by = c("geo", "area_of_residence", "date")) %>%
  mutate(gap_v = v - v_sum,
         gap_s = s_forecast - s_sum)
```

#sum of overseas and USA, and compare total international
```{r}
check_s_3_3 <- final_df_s %>%
  filter(area_of_residence %in% c("Residents of countries other than the United States of America", "United States of America residents, tourists (overnight)")) %>%
  group_by(geo, date) %>%
  summarise(v_sum = sum(v, na.rm = TRUE),
            s_sum = sum(s_forecast, na.rm = TRUE), .groups = "drop") %>%
  mutate(area_of_residence = "Non-resident visitors to Canada") %>%
  ungroup() %>%
  left_join(select(final_df_s, geo, area_of_residence, date, v, s_forecast), by = c("geo", "area_of_residence", "date")) %>%
  mutate(gap_v = v - v_sum,
         gap_s = s_forecast - s_sum)
```

#check negative values
```{r}
check_v_4 <- final_df_v_process1 %>%
  mutate(v_neg = ifelse(value < 0, 1, 0))

check_s_4 <- final_df_s %>%
  mutate(v_neg = ifelse(v < 0, 1, 0),
         s_neg = ifelse(s_forecast < 0, 1, 0))
```

#check for negative yoy growth
```{r}
check_v_5 <- final_df_v_process1 %>%
  # select(loc_ttl, date, orig_ttl, v_est, s_est) %>%
  group_by(geo, country_of_residence) %>%
  mutate(prev_v = lag(value)) %>%
  mutate(growth_v = value/prev_v) %>%
  mutate(check_v = case_when(growth_v < 1 ~ 1,
                             TRUE ~ 0)) %>%
  ungroup() %>%
  filter(date >= "2024-01-01")

check_s_5 <- final_df_s %>%
  # select(loc_ttl, date, orig_ttl, v_est, s_est) %>%
  group_by(geo, area_of_residence) %>%
  mutate(prev_v = lag(v),
         prev_s = lag(s_forecast)) %>%
  mutate(growth_v = v/prev_v,
         growth_s = s_forecast/prev_s) %>%
  mutate(check_v = case_when(growth_v < 1 ~ 1,
                             TRUE ~ 0),
         check_s = case_when(growth_s < 1 ~ 1,
                             TRUE ~ 0)) %>%
  ungroup() %>%
  filter(date >= "2024-01-01")
```

#if v from final_v the same as v from final_s
```{r}
check_6 <- final_df_v_process1 %>%
  filter(country_of_residence %in% c("United States of America residents entering Canada", "Residents of countries other than the United States of America entering Canada")) %>%
  mutate(ttl = case_when(country_of_residence == "United States of America residents entering Canada" ~ "United States of America residents, tourists (overnight)",
                         country_of_residence == "Residents of countries other than the United States of America entering Canada" ~ "Residents of countries other than the United States of America",
                         TRUE ~ NA)) %>%
  left_join(final_df_s, by = c("geo", "ttl" = "area_of_residence", "date")) %>%
  mutate(gap_v = v-value) %>%
  filter(date >= "2018-01-01")
```


#For comparison file
#gts for visits dataset
```{r}
gts_v <- gts_current_ctry_NES %>%
  filter(variable == "v" ) %>%
  mutate(value = 1000*value)

gts_ctry <- gts_v %>%
  filter(orig_code %in% unique(ctry_tour$orig_code))

gts_ovs <- gts_v %>%
  filter(orig_ttl != "United States") %>%
  group_by(loc_ttl, loc_code, date, variable) %>%
  summarise(value = sum(value, na.rm = TRUE), .groups = "drop") %>%
  ungroup() %>%
  mutate(orig_ttl = "Overseas", 
         orig_code = "OVS")

gts_tot <- gts_v %>%
  group_by(loc_ttl, loc_code, date, variable) %>%
  summarise(value = sum(value, na.rm = TRUE), .groups = "drop") %>%
  ungroup() %>%
  mutate(orig_ttl = "Total international", 
         orig_code = "TOT")

gts_v_complete <- gts_ctry %>%
  rbind(gts_ovs, gts_tot)

# gts_v_rest_reg <- gts_v %>%
#   left_join(ctry_def2, by = c("orig_code")) %>%
#   filter(!orig_code %in% unique(ctry_tour$orig_code)) %>%
#   group_by(date, low_lev_reg) %>%
#   summarise(sum_rest_reg = sum(value, na.rm = TRUE), .groups = "drop") %>%
#   ungroup() %>%
#   group_by(low_lev_reg) %>%
#   mutate(growth = sum_rest_reg/sum_rest_reg[date == "2023-01-01"]) %>%
#   ungroup()

```

#gts for spending dataset
```{r}
gts_s <- gts_current_ctry_NES %>%
  filter(variable == "s") %>%
  select(loc_code, date, orig_code, s = value)

gts_spend_per <- gts_s %>%
  filter(orig_code != "CAN") %>%
  left_join(gts_v, by = c("loc_code", "date", "orig_code")) %>%
  left_join(ctry_def3, by = c("orig_code")) %>%
  group_by(loc_code, date, s_ttl) %>%
  summarise(s = sum(s*1000000, na.rm = TRUE),
            v= sum(value, na.rm = TRUE),
            .groups =  "drop") %>%
  ungroup() %>%
  mutate(gts_spend_per = s/v*1000,
         gts_spend_per = ifelse(is.na(gts_spend_per) | is.infinite(gts_spend_per), 0, gts_spend_per)) %>%
  select(date, s_ttl, v, s, gts_spend_per)

gts_spend_per_ovs <- gts_spend_per %>%
  filter(s_ttl != "United States of America residents, tourists (overnight)") %>%
  group_by(date) %>%
  summarise(v = sum(v, na.rm = TRUE),
            s = sum(s, na.rm = TRUE),
            .groups = "drop") %>%
  ungroup() %>%
  mutate(gts_spend_per = s/v*1000,
         gts_spend_per = ifelse(is.na(gts_spend_per) | is.infinite(gts_spend_per), 0, gts_spend_per)) %>%
  mutate(s_ttl = "Overseas")

gts_spend_per_tot <- gts_spend_per %>%
  group_by(date) %>%
  summarise(v = sum(v, na.rm = TRUE),
            s = sum(s, na.rm = TRUE),
            .groups = "drop") %>%
  ungroup() %>%
  mutate(gts_spend_per = s/v*1000,
         gts_spend_per = ifelse(is.na(gts_spend_per) | is.infinite(gts_spend_per), 0, gts_spend_per)) %>%
  mutate(s_ttl = "Total international")

gts_s_complete <- gts_spend_per %>%
  rbind(gts_spend_per_ovs, gts_spend_per_tot)

# #use Netherlands to fill Belgium
# gts_bel <- gts_spend_per %>%
#   filter(s_ttl == "Netherlands") %>%
#   mutate(s_ttl = "Belgium")
# 
# gts_spend_per_complete <- gts_spend_per %>%
#   filter(s_ttl != "Belgium") %>%
#   rbind(gts_bel)
```



#dataset for client
```{r}
final_v <- final_df_v_process1 %>%
  rename(v = value) %>%
  mutate(country_of_residence = ifelse(country_of_residence == "Türkiye", "Turkiye", country_of_residence))

final_s <- final_df_s %>%
  rename(country_of_residence = area_of_residence, 
         s = s_forecast,
         spend_per = revised_spend_per) %>%
  mutate(spend_per = s/v *1000) %>%
  mutate(spend_per = ifelse(is.na(spend_per) | is.infinite(spend_per), 0, spend_per))
```

```{r}
v_ref <- v_ref_for_client
s_ref <- reg_def2 %>%
  rename(country_of_residence = s_high_lev_reg)
```

```{r}
write_rds(final_v, here(path(path_te_shared, path_output),"deliver", "final_v.rds"))
write.csv(final_v, here(path(path_te_shared, path_output),"deliver", "final_v.csv"))

write_rds(final_s, here(path(path_te_shared, path_output),"deliver", "final_s.rds"))
write.csv(final_s, here(path(path_te_shared, path_output),"deliver", "final_s.csv"))

write.csv(v_ref, here(path(path_te_shared, path_output),"deliver", "v_ref.csv"))
write.csv(s_ref, here(path(path_te_shared, path_output),"deliver", "s_ref.csv"))


write.csv(gts_v_complete, here(path(path_te_shared, path_output), "gts_v_complete.csv"))
write.csv(gts_s_complete, here(path(path_te_shared, path_output), "gts_s_complete.csv"))
```

