---
title: "R Notebook"
output: html_notebook
---

```{r}
library(dplyr)
library(knitr)
library(tidyverse)
library(stringr)
library(here)
library(fs)
library(readxl)

source(here("scripts", "set_vars.R"))
options(scipen=999)
```

```{r}
final_v<- read_rds(here(path(path_te_shared, path_output),"deliver", "final_v.rds"))
final_s<- read_rds(here(path(path_te_shared, path_output),"deliver", "final_s.rds"))

#archived
final_v0<- read_rds(here(path(path_te_shared, path_output),"deliver/archived", "final_v0.rds"))
final_s0<- read_rds(here(path(path_te_shared, path_output),"deliver/archived", "final_s0.rds"))

#reference
ctry_def2<- read_rds(here("output_data", "ctry_def2.rds"))
ctry_def3<- read_rds(here("output_data", "ctry_def3.rds"))
reg_def <- read_rds(here("output_data", "reg_def.rds"))
```

#check diff visits
```{r}
check_v <- final_v %>%
  left_join(final_v0, by = c("geo", "country_of_residence", "orig_code", "low_lev_reg", "high_lev_reg", "date")) %>%
  mutate(gap_v = v.x/v.y)
```
#check diff spending
```{r}
check_s <- final_s %>%
  left_join(final_s0, by = c("geo", "country_of_residence", "date")) %>%
  mutate(gap_v = v.x-v.y,
         gap_s = s.x-s.y)
```


#check sum of provinces and total Canada
```{r}
check_v_1 <- final_v %>%
  filter(geo != "Canada") %>%
  group_by(country_of_residence, date) %>%
  summarise(sum_v = sum(v, na.rm = TRUE), .groups = "drop") %>%
  ungroup() %>%
  left_join(final_v %>%filter(geo == "Canada"), by = c("country_of_residence", "date")) %>%
  mutate(gap_v = v - sum_v)
```

```{r}
check_s_1 <- final_s %>%
  filter(geo != "Canada") %>%
  group_by(country_of_residence, date) %>%
  summarise(sum_v = sum(v, na.rm = TRUE),
            sum_s = sum(s, na.rm = TRUE),
            .groups = "drop") %>%
  ungroup() %>%
  left_join(final_s %>%filter(geo == "Canada"), by = c("country_of_residence", "date")) %>%
  mutate(gap_v = v - sum_v,
         gap_s = s - sum_s)
```




#check sum of country and rest of region, and compare with low lev region
```{r}
check_v_3_1 <- final_v %>%
  filter(!(country_of_residence %in% unique(reg_def$low_lev_reg)) & !is.na(low_lev_reg)) %>%
  group_by(geo, low_lev_reg, date) %>%
  summarise(v_sum = sum(v, na.rm = TRUE), .groups = "drop") %>%
  ungroup() %>%
  left_join(select(final_v, geo, country_of_residence, date, v), by = c("geo", "low_lev_reg" = "country_of_residence", "date")) %>%
  mutate(gap_v = v - v_sum)
```

#check sum of low lev reg, and compare with high lev region
```{r}
check_v_3_2 <- check_v_3_1 %>%
  select(-c(v, gap_v)) %>%
  left_join(reg_def, by = c("low_lev_reg")) %>%
  group_by(geo, high_lev_reg, date) %>%
  summarise(v_sum = sum(v_sum, na.rm = TRUE), .groups = "drop") %>%
  ungroup() %>%
  left_join(select(final_v, geo, country_of_residence, date, v), by = c("geo", "high_lev_reg" = "country_of_residence", "date")) %>%
  mutate(gap_v = v - v_sum)
```

#check sum of high lev reg, and compare with overseas
```{r}
check_v_3_3 <- check_v_3_2 %>%
  select(-c(v, high_lev_reg, gap_v)) %>%
  group_by(geo, date) %>%
  summarise(v_sum = sum(v_sum, na.rm = TRUE), .groups = "drop") %>%
  ungroup() %>%
  mutate(country_of_residence = "Residents of countries other than the United States of America entering Canada") %>%
  left_join(select(final_v, geo, country_of_residence, date, v), by = c("geo", "country_of_residence", "date")) %>%
  mutate(gap_v = v - v_sum)
```

#check sum of overseas and USA, and compare with total international
```{r}
check_v_3_4 <- final_v %>%
  filter(country_of_residence %in% c("Residents of countries other than the United States of America entering Canada", "United States of America residents entering Canada")) %>%
  group_by(geo, date) %>%
  summarise(v_sum = sum(v, na.rm = TRUE), .groups = "drop") %>%
  ungroup() %>%
  mutate(country_of_residence = "Non-resident visitors entering Canada") %>%
  left_join(select(final_v, geo, country_of_residence, date, v), by = c("geo", "country_of_residence", "date")) %>%
  mutate(gap_v = v - v_sum)
```


#check sum of things in spending

```{r}
reg_def2 <- data.frame(s_ttl= c("Not elsewhere specified-country in Americas, countries other than the United States of America", "Americas, countries other than the United States of America - other countries", "Mexico", "Caribbean", "Brazil", "Europe - other countries", "Belgium", "France", "Germany", "Netherlands", "Switzerland", "Scandinavia", "United Kingdom", "Italy", "Not elsewhere specified-country in Europe", "Africa - other countries", "Northern Africa", "Not elsewhere specified-country in Africa", "Asia and Oceania - other countries", "China", "Hong Kong", "Japan", "Southern Asia", "India", "Australia", "New Zealand","Not elsewhere specified-country in Asia and Oceania", "Antarctica and Adjacent Islands - other countries", "Not elsewhere specified-country"),
                             s_high_lev_reg=c("Americas, countries other than the United States of America", "Americas, countries other than the United States of America", "Americas, countries other than the United States of America", "Americas, countries other than the United States of America", "Americas, countries other than the United States of America", "Europe", "Europe", "Europe", "Europe", "Europe", "Europe", "Europe", "Europe", "Europe", "Europe", "Africa", "Africa", "Africa", "Asia and Oceania", "Asia and Oceania", "Asia and Oceania", "Asia and Oceania", "Asia and Oceania", "Asia and Oceania", "Asia and Oceania", "Asia and Oceania", "Asia and Oceania", "Antarctica and Adjacent Islands", "Not elsewhere specified-region"))
```

#sum of country and rest of region, and compare with region
```{r}
check_s_3_1 <- final_s %>%
  left_join(reg_def2, by = c("country_of_residence" = "s_ttl")) %>%
  filter(!is.na(s_high_lev_reg)) %>%
  group_by(geo, s_high_lev_reg, date) %>%
  summarise(v_sum = sum(v, na.rm = TRUE),
            s_sum = sum(s, na.rm = TRUE), .groups = "drop") %>%
  ungroup() %>%
  left_join(select(final_s, geo, country_of_residence, date, v, s), by = c("geo", "s_high_lev_reg" = "country_of_residence", "date")) %>%
  mutate(gap_v = v - v_sum,
         gap_s = s - s_sum)
```

#sum of region, and compare with overseas
```{r}
check_s_3_2 <- check_s_3_1 %>%
  group_by(geo, date) %>%
  summarise(v_sum = sum(v_sum, na.rm = TRUE),
            s_sum = sum(s_sum, na.rm = TRUE), .groups = "drop") %>%
  mutate(country_of_residence = "Residents of countries other than the United States of America") %>%
  ungroup() %>%
  left_join(select(final_s, geo, country_of_residence, date, v, s), by = c("geo", "country_of_residence", "date")) %>%
  mutate(gap_v = v - v_sum,
         gap_s = s - s_sum)
```

#sum of overseas and USA, and compare total international
```{r}
check_s_3_3 <- final_s %>%
  filter(country_of_residence %in% c("Residents of countries other than the United States of America", "United States of America residents, tourists (overnight)")) %>%
  group_by(geo, date) %>%
  summarise(v_sum = sum(v, na.rm = TRUE),
            s_sum = sum(s, na.rm = TRUE), .groups = "drop") %>%
  mutate(country_of_residence = "Non-resident visitors to Canada") %>%
  ungroup() %>%
  left_join(select(final_s, geo, country_of_residence, date, v, s), by = c("geo", "country_of_residence", "date")) %>%
  mutate(gap_v = v - v_sum,
         gap_s = s - s_sum)
```

#check negative values
```{r}
check_v_4 <- final_v %>%
  mutate(v_neg = ifelse(v < 0, 1, 0))

check_s_4 <- final_s %>%
  mutate(v_neg = ifelse(v < 0, 1, 0),
         s_neg = ifelse(s < 0, 1, 0))
```

#check for negative yoy growth
```{r}
check_v_5 <- final_v %>%
  group_by(geo, country_of_residence) %>%
  mutate(prev_v = lag(v)) %>%
  mutate(growth_v = v/prev_v) %>%
  mutate(check_v = case_when(growth_v < 1 ~ 1,
                             TRUE ~ 0)) %>%
  ungroup() %>%
  filter(date >= "2024-01-01")

check_s_5 <- final_s %>%
  group_by(geo, country_of_residence) %>%
  mutate(prev_v = lag(v),
         prev_s = lag(s)) %>%
  mutate(growth_v = v/prev_v,
         growth_s = s/prev_s) %>%
  mutate(check_v = case_when(growth_v < 1 ~ 1,
                             TRUE ~ 0),
         check_s = case_when(growth_s < 1 ~ 1,
                             TRUE ~ 0)) %>%
  ungroup() %>%
  filter(date >= "2024-01-01")
```

#if v from final_v the same as v from final_s
```{r}
check_6 <- final_v %>%
  filter(country_of_residence %in% c("United States of America residents entering Canada", "Residents of countries other than the United States of America entering Canada")) %>%
  mutate(ttl = case_when(country_of_residence == "United States of America residents entering Canada" ~ "United States of America residents, tourists (overnight)",
                         country_of_residence == "Residents of countries other than the United States of America entering Canada" ~ "Residents of countries other than the United States of America",
                         TRUE ~ NA)) %>%
  left_join(final_s, by = c("geo", "ttl" = "country_of_residence", "date")) %>%
  mutate(gap_v = v.x-v.y) %>%
  filter(date >= "2018-01-01")
```





